---
title: "Pangenome compendia"
output: powerpoint_presentation
date: "2025-08-06"
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  #dev = "#png",
  dpi = 300,
  fig.width = 5,
  fig.height = 4
)
```

```{r, warning=F, message=F, include=F}
library(librarian)
shelf(KEGGREST, 
      EnrichmentBrowser,
      ComplexHeatmap,
      circlize,
      RColorBrewer,
      DESeq2,
      dplyr,
      corrplot,
      ggplot2,
      stringr,
      reshape2,
      ggprism,
      magick,
      readr)
#library(KEGGREST)
#library(EnrichmentBrowser)
#library(ComplexHeatmap)
#library(circlize)
#library(RColorBrewer)
#library(DESeq2)
#library(dplyr)
#library(corrplot)
#library(ggplot2)
```



## Load data

```{r}
se_counts <- read.csv('num_reads_sepi_pan_genome_sepi_comp_k15.csv', stringsAsFactors = F)
sa_counts <- read.csv('num_reads_saur_pan_genome_saur_comp_k15.csv', stringsAsFactors = F)
sa_cross_counts <- read.csv('num_reads_sepi_pan_genome_saur_comp_k15.csv', stringsAsFactors = F)

```

## Load Run Tables

```{r}
se_run_table <- read.csv('SraRunTable_Se_rna_2024_02_07_st_extra.csv', stringsAsFactors = F)
sa_run_table <- read.csv('SraRunTable_Sa_rna_2024_02_07_st_extra.csv', stringsAsFactors = F)

```



## Data frames


```{r}
se_studies <- sapply(colnames(se_counts)[-c(1,2)], function(x) 
  se_run_table$SRA.Study[se_run_table$Experiment == x][1])
se_strains <- sapply(colnames(se_counts)[-c(1,2)], function(x) 
  se_run_table$str_detect_group[se_run_table$Experiment == x][1])

###

sa_studies <- sapply(colnames(sa_counts)[-c(1,2)], function(x) 
  sa_run_table$SRA.Study[sa_run_table$Experiment == x][1])
sa_strains <- sapply(colnames(sa_counts)[-c(1,2)], function(x) 
  sa_run_table$str_detect_group[sa_run_table$Experiment == x][1])

## sa_cross_counts
sa_cross_studies <- sapply(colnames(sa_cross_counts)[-c(1,2)], function(x) 
  sa_run_table$SRA.Study[sa_run_table$Experiment == x][1])
sa_cross_strains <- sapply(colnames(sa_cross_counts)[-c(1,2)], function(x) 
  sa_run_table$str_detect_group[sa_run_table$Experiment == x][1])


```



## Se initital heatmap - trimmed for plotting



```{r, warning=F, message=F,, fig.width=12, fig.height=6}
se_counts_t <- se_counts[,-c(1,2)]
se_counts_t <- se_counts_t[rowSums(se_counts_t) > 50000,]

se_log_counts <- log(se_counts_t + 1)


se_col_fun = colorRamp2(c(min(se_log_counts), median(data.matrix(se_log_counts)), max(se_log_counts)), c("grey", "#E7D4E8", "#762A83"))

se_ann_colors <- rep(brewer.pal(6,'Greys'),15)[1:length(unique(se_studies))]

names(se_ann_colors) <- unique(se_studies)

se_st_colors <- rep(brewer.pal(8,'Dark2'),3)[1:length(unique(se_strains))]
names(se_st_colors) <- unique(se_strains)

ha <- HeatmapAnnotation(study = se_studies, 
                        strain = se_strains,
                        show_legend = c(F,T), 
                        col = list( study = se_ann_colors, 
                                                     strain = se_st_colors))

h <- Heatmap(se_log_counts, 
             col = se_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             show_row_names = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(2,"in"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=0),
             )
h

#png('plots/Se_comp_trimmed.#png', height=6, width=12, unit = "in", res = 150)
#h
#dev.off()

#pdf('plots/se_comp_genes_se16_mapped_no_spo_dups_small.pdf', height=3, width=6)
#h
#dev.off()
```

## Se trim based on pangenome

```{r}
se_gene_pa <- read.csv('sepi_out_1707234787/gene_presence_absence.csv', stringsAsFactors = F)
se_gene_anns <- read.csv('sepi_pan_genome/sepi_pan_genome.tsv', sep = '\t', stringsAsFactors = F)
se_gene_loci <- read.table('sepi_pan_genome/sepi_pan_genome_feature_table.tsv', stringsAsFactors = F,
                           header = F, skip = 1, sep = ">")
se_gene_loci$locus_id <- str_extract(se_gene_loci$V2, '[a-zA-Z0-9_]+')
se_gene_loci$name_id <- str_extract(se_gene_loci$V2, 'GOBBEKFC_[0-9]*')

se_gene_pa$compound_name <- apply(se_gene_pa[,c(15:20)], 1, FUN = function(x) paste(x, collapse  = " ") )
se_gene_pa$first_name <- sapply(se_gene_pa$compound_name, function(x) str_extract(x, "[0-9a-zA-Z_]+"))
sum(se_gene_pa$first_name %in% se_counts$X)

```



```{r, fig.height=3, fig.width=3}
se_mu <- rowMeans(se_counts[,-c(1,2)])
names(se_mu) <- se_counts$X



se_gene_pa$mean_exp <- se_mu[se_gene_pa$first_name]

g <- ggplot(se_gene_pa, aes(y = mean_exp, x = factor(No..isolates))) +
  geom_boxplot() +
  scale_y_log10() +
  theme_prism()
g
```

## Mean exp table

```{r}
table(se_gene_pa$No..isolates, se_gene_pa$mean_exp == 0)
```

## Zeros

```{r, fig.height=3, fig.width=3}
se_zs <- rowSums(se_counts[,-c(1,2)] == 0)
names(se_zs) <- se_counts$X
se_gene_pa$zeros <- se_zs[se_gene_pa$first_name]


g <- ggplot(se_gene_pa, aes(y = zeros, x = factor(No..isolates))) +
  geom_boxplot() +
  #scale_y_log10() +
  theme_prism()
g
```


## Heatmap


```{r, warning=F, message=F,, fig.width=12, fig.height=6}
rownames(se_gene_pa) <- se_gene_pa$first_name
se_counts_t <- se_counts[se_gene_pa[se_counts$X,"No..isolates"] > 1,-c(1,2)]
#se_counts_t <- se_counts_t[rowSums(se_counts_t) > 50000,]

se_log_counts <- log(se_counts_t + 1)


se_col_fun = colorRamp2(c(min(se_log_counts), median(data.matrix(se_log_counts)), max(se_log_counts)), c("grey", "#E7D4E8", "#762A83"))

se_ann_colors <- rep(brewer.pal(6,'Greys'),15)[1:length(unique(se_studies))]

names(se_ann_colors) <- unique(se_studies)

se_st_colors <- rep(brewer.pal(8,'Dark2'),3)[1:length(unique(se_strains))]
names(se_st_colors) <- unique(se_strains)

ha <- HeatmapAnnotation(study = se_studies, 
                        strain = se_strains,
                        show_legend = c(F,T), 
                        col = list( study = se_ann_colors, 
                                                     strain = se_st_colors))

h <- Heatmap(se_log_counts, 
             col = se_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             show_row_names = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(2,"in"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=0),
             )
h

#png('plots/Se_comp_trimmed.#png', height=6, width=12, unit = "in", res = 150)
#h
#dev.off()

#pdf('plots/se_comp_genes_se16_mapped_no_spo_dups_small.pdf', height=3, width=6)
#h
#dev.off()
```



## Sa initital heatmap - trimmed for plotting


```{r, warning=F, message=F,, fig.width=12, fig.height=6}
sa_counts_t <- sa_counts[,-c(1,2)]
rownames(sa_counts_t) <- sa_counts$X
sa_counts_t <- sa_counts_t[rowSums(sa_counts[,-c(1,2)]) > 500000,]

sa_log_counts <- log(sa_counts_t + 1)


sa_col_fun = colorRamp2(c(min(sa_log_counts), median(data.matrix(sa_log_counts)), max(sa_log_counts)), c("grey", "#E7D4E8", "#762A83"))


sa_ann_colors <- rep(brewer.pal(6,'Greys'),100)[1:length(unique(sa_studies))]

names(sa_ann_colors) <- unique(sa_studies)

sa_st_colors <- rep(brewer.pal(8,'Dark2'),3)[1:length(unique(sa_strains))]
names(sa_st_colors) <- unique(sa_strains)

ha <- HeatmapAnnotation(study = sa_studies, 
                        strain = sa_strains,
                        show_legend = c(F,T), 
                        col = list( study = sa_ann_colors, 
                                                     strain = sa_st_colors))

h <- Heatmap(sa_log_counts, 
             col = sa_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             show_row_names = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(2,"in"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=0),
             )
h

#png('plots/Sa_comp_trimmed.#png', height=6, width=24, unit = "in", res = 150)
#h
#dev.off()
#h
#pdf('plots/se_comp_genes_se16_mapped_no_spo_dups_small.pdf', height=3, width=6)
#h
#dev.off()
```

## Sa trim based on pangenome

```{r}
###


sa_gene_pa <- read.csv('saur_out_1707260617/gene_presence_absence.csv', stringsAsFactors = F)
sa_gene_anns <- read.csv('saur_pan_genome/saur_pan_genome.tsv', sep = '\t', stringsAsFactors = F)
sa_gene_loci <- read.table('saur_pan_genome/saur_pan_genome_feature_table.tsv', stringsAsFactors = F,
                           header = F, skip = 1, sep = ">")
sa_gene_loci$locus_id <- str_extract(sa_gene_loci$V2, '[a-zA-Z0-9_]+')
sa_gene_loci$name_id <- str_extract(sa_gene_loci$V2, 'AADLDMHG_[0-9]*')

sa_gene_pa$compound_name <- apply(sa_gene_pa[,c(15:32)], 1, FUN = function(x) paste(x, collapse  = " ") )
sa_gene_pa$first_name <- sapply(sa_gene_pa$compound_name, function(x) str_extract(x, "[0-9a-zA-Z_]+"))
sum(sa_gene_pa$first_name %in% sa_counts$X)
```


```{r, fig.height=3, fig.width=3}
sa_mu <- rowMeans(sa_counts[,-c(1,2)])
names(sa_mu) <- sa_counts$X



sa_gene_pa$mean_exp <- sa_mu[sa_gene_pa$first_name]

g <- ggplot(sa_gene_pa, aes(y = mean_exp, x = factor(No..isolates))) +
  geom_boxplot() +
  scale_y_log10() +
  theme_prism()
g
```

## Table


```{r}
table(sa_gene_pa$No..isolates, sa_gene_pa$mean_exp == 0)
```


## Zeros plot


```{r, fig.height=3, fig.width=4}
sa_zs <- rowSums(sa_counts[,-c(1,2)] == 0)
names(sa_zs) <- sa_counts$X
sa_gene_pa$zeros <- sa_zs[sa_gene_pa$first_name]


g <- ggplot(sa_gene_pa, aes(y = zeros, x = factor(No..isolates))) +
  geom_boxplot() +
  #scale_y_log10() +
  theme_prism()
g
```


## Heatmap


```{r, warning=F, message=F,, fig.width=12, fig.height=6}
rownames(sa_gene_pa) <- make.unique(sa_gene_pa$first_name)
sa_counts_t <- sa_counts[sa_gene_pa[sa_counts$X[sa_counts$X %in% sa_gene_pa$first_name],"No..isolates"] > 1,-c(1,2)]

sa_log_counts <- log(sa_counts_t + 1)


sa_col_fun = colorRamp2(c(min(sa_log_counts), median(data.matrix(sa_log_counts)), max(sa_log_counts)), c("grey", "#E7D4E8", "#762A83"))


sa_ann_colors <- rep(brewer.pal(6,'Greys'),100)[1:length(unique(sa_studies))]

names(sa_ann_colors) <- unique(sa_studies)

sa_st_colors <- rep(brewer.pal(8,'Dark2'),3)[1:length(unique(sa_strains))]
names(sa_st_colors) <- unique(sa_strains)

ha <- HeatmapAnnotation(study = sa_studies, 
                        strain = sa_strains,
                        show_legend = c(F,T), 
                        col = list( study = sa_ann_colors, 
                                                     strain = sa_st_colors))

h <- Heatmap(sa_log_counts, 
             col = sa_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             show_row_names = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(2,"in"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=0),
             )
h

#png('plots/Sa_comp_trimmed.#png', height=6, width=24, unit = "in", res = 150)
#h
#dev.off()
#h
#pdf('plots/se_comp_genes_se16_mapped_no_spo_dups_small.pdf', height=3, width=6)
#h
#dev.off()


## Sa - cross mapped initital heatmap - trimmed for plotting
```


## Sa cross counts trim for plotting

```{r, warning=F, message=F,, fig.width=12, fig.height=6}
sa_cross_counts_t <- sa_cross_counts[,-c(1,2)]
sa_cross_counts_t <- sa_cross_counts_t[rowSums(sa_cross_counts[,-c(1,2)]) > 50000,]

sa_cross_log_counts <- log(sa_cross_counts_t + 1)


sax_col_fun = colorRamp2(c(min(sa_cross_log_counts), median(data.matrix(sa_cross_log_counts)), max(sa_cross_log_counts)), c("grey", "#E7D4E8", "#762A83"))



sax_ann_colors <- rep(brewer.pal(6,'Greys'),100)[1:length(unique(sa_cross_studies))]

names(sax_ann_colors) <- unique(sa_cross_studies)

sax_st_colors <- rep(brewer.pal(8,'Dark2'),3)[1:length(unique(sa_cross_strains))]
names(sax_st_colors) <- unique(sa_cross_strains)

ha <- HeatmapAnnotation(study = sa_cross_studies, 
                        strain = sa_cross_strains,
                        show_legend = c(F,T), 
                        col = list( study = sax_ann_colors, 
                                                     strain = sax_st_colors))




h <- Heatmap(sa_cross_log_counts, 
             col = sax_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             show_row_names = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=0),
             )
h

#png('plots/Sa_cross_mapped_comp_trimmed_noclust.#png', height=6, width=24, unit = "in", res = 150)
#h
#dev.off()
#h
#pdf('plots/se_comp_genes_se16_mapped_no_spo_dups_small.pdf', height=3, width=6)
#h
#dev.off()
```


## Sa cross trim by pangenome

```{r}
sum(se_gene_pa$first_name %in% sa_cross_counts$X)
```


```{r, fig.height=3, fig.width=3}
sax_mu <- rowMeans(sa_cross_counts[,-c(1,2)])
names(sax_mu) <- sa_cross_counts$X



se_gene_pa$cross_mean_exp <- sax_mu[se_gene_pa$first_name]

g <- ggplot(se_gene_pa, aes(y = cross_mean_exp, x = factor(No..isolates))) +
  geom_boxplot() +
  scale_y_log10() +
  theme_prism()
g
```


## Check Zero counts



```{r}
table(se_gene_pa$No..isolates, se_gene_pa$cross_mean_exp == 0)
```


## Plot zeros


```{r, fig.height=3, fig.width=3}
sax_zs <- rowSums(sa_cross_counts[,-c(1,2)] == 0)
names(sax_zs) <- sa_cross_counts$X
se_gene_pa$cross_zeros <- sax_zs[se_gene_pa$first_name]


g <- ggplot(se_gene_pa, aes(y = cross_zeros, x = factor(No..isolates))) +
  geom_boxplot() +
  #scale_y_log10() +
  theme_prism()
g
```

## Heatmap


```{r, warning=F, message=F,, fig.width=12, fig.height=6}
rownames(sa_cross_counts) <- sa_cross_counts$X
sa_cross_counts_t <- sa_cross_counts_t[se_gene_pa[se_counts$X,"No..isolates"] > 1,-c(1,2)]

sa_cross_log_counts <- log(sa_cross_counts_t + 1)


sax_col_fun = colorRamp2(c(min(sa_cross_log_counts), median(data.matrix(sa_cross_log_counts)), max(sa_cross_log_counts)), c("grey", "#E7D4E8", "#762A83"))



sax_ann_colors <- rep(brewer.pal(6,'Greys'),100)[1:length(unique(sa_cross_studies))]

names(sax_ann_colors) <- unique(sa_cross_studies)

sax_st_colors <- rep(brewer.pal(8,'Dark2'),3)[1:length(unique(sa_cross_strains))]
names(sax_st_colors) <- unique(sa_cross_strains)

ha <- HeatmapAnnotation(study = sa_cross_studies, 
                        strain = sa_cross_strains,
                        show_legend = c(F,T), 
                        col = list( study = sax_ann_colors, 
                                                     strain = sax_st_colors))




h <- Heatmap(sa_cross_log_counts, 
             col = sax_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             show_row_names = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=0),
             )
h

#png('plots/Sa_cross_mapped_comp_trimmed_noclust.#png', height=6, width=24, unit = "in", res = 150)
#h
#dev.off()
#h
#pdf('plots/se_comp_genes_se16_mapped_no_spo_dups_small.pdf', height=3, width=6)
#h
#dev.off()
```

## HK genes

From  paper, hk genes:
* gmk (SE_0885) - got from feature table above
* DHFR genes dfrc (SE_1119) and thyF (SE_1120) (megablast emb|Z48233.1|)
* tpi (SE_0560) (megablstgb|KY398645.1|)
* 16S rRNA (SEr06) - had to blast ATCC 14990 against S.e. ATCC 122238 (megablast dbj|D83363.1|)
* hsp-60  groL (SE_1629)(megablast gb|AF029245.1|)






## Se HK gene dist


```{r, warning=F, message=F,, fig.width=4, fig.height=4}
se_hk_genes <- se_gene_pa[sapply(se_gene_pa$Gene, 
                                             function(x){
                                               y <- str_detect(x, 
                                                               c("gmk","thyF","tpi","gyrB",
                                                                 "groL","hsp","dfr","16S",
                                                                 "tpi","groE","rsm", "rpoD",
                                                                 "dna"))
                                               TRUE %in% y
                                               }),]

se_hk_genes_names <- se_hk_genes$Gene[se_hk_genes$No..isolates == 6]
se_hk_genes_loci <- se_hk_genes$GCF_000007645.1[se_hk_genes$No..isolates == 6]

se_hk_loci <- se_hk_genes_loci[se_hk_genes_loci %in% se_counts$X]


se_hk_counts  <- se_counts[se_counts$X %in% se_hk_loci,]

hk_probs <- quantile(apply(se_hk_counts[,-c(1,2)],2,FUN = 
                            function(x) log(max(median(x),1))), 
                    probs = c(.05,.995))

#pdf('plots/se16_comp_hk_counts.pdf', height=4, width=4)
plot(hist(apply(se_hk_counts[,-c(1,2)],2,FUN = function(x) log(max(median(x),1))), 
          breaks = 75, plot=F),
     main = 'HK genes', xlab = 'log counts (floor 1)')
abline(v = hk_probs[1], col='red',lty=2)
abline(v = hk_probs[2], col='red',lty=2)
#dev.off()

```


## Se zero-count genes dist

```{r, warning=F, message=F,, fig.width=4, fig.height=4}
zc_probs <- quantile(apply(se_counts[,-c(1,2)],2,FUN = function(x) log(max(sum(x==0),1))), 
                    probs = c(.032,.968))
#pdf('plots/se16_comp_zero_counts.pdf', height=4, width=4)
plot(hist(apply(se_counts[,-c(1,2)],2,FUN = function(x) log(max(sum(x==0),1))),
          breaks = 50, plot=F),
     main = 'zero count genes', xlab = 'log zero counts(floor 1)')
abline(v = zc_probs[1], col='red',lty=2)
abline(v = zc_probs[2], col='red',lty=2)
#dev.off()
```




## Filter Se data


```{r}
hk_medians <- apply(se_hk_counts[,-c(1,2)],2,FUN = 
                            function(x) log(max(median(x),1)))
zero_counts <- apply(se_counts[,-c(1,2)],2,FUN = 
                       function(x) log(max(sum(x==0),1)))

hk_lims <- hk_probs
zc_lims <- zc_probs
se_counts_filtered <- sapply(c(1:length(hk_medians)), 
                             function(x){
                               if(hk_medians[x] > hk_lims[1] & 
                                  #hk_medians[x] < hk_lims[2] &
                                  #zero_counts[x] > zc_lims[1] &
                                  zero_counts[x] < zc_lims[2]){T}else{F}
                             })

```


## Heatmap of Se data showing filter


```{r, warning=F, message=F,, fig.width=12, fig.height=6}


ha <- HeatmapAnnotation(study = se_studies, 
                        strain = se_strains,
                        filter = se_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = se_ann_colors,
                                    strain = se_st_colors,
                                    filter = c('TRUE'="white",'FALSE'="black")))


h <- Heatmap(se_log_counts, 
             col = se_col_fun, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
```


## Heatmap of Se data filtered



```{r, warning=F, message=F,, fig.width=12, fig.height=6}

ha <- HeatmapAnnotation(study = se_studies[se_counts_filtered], 
                        strain = se_strains[se_counts_filtered],
                        #filter = se_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = se_ann_colors,
                                    strain = se_st_colors#,
                                    #filter = c('TRUE'="white",'FALSE'="black")
                                    ))

h <- Heatmap(se_log_counts[,se_counts_filtered], 
             col = se_col_fun, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
```


## Sa HK gene dists

```{r, warning=F, message=F,, fig.width=4, fig.height=4}
sa_hk_genes <- sa_gene_pa[sapply(sa_gene_pa$Gene, 
                                             function(x){
                                               y <- str_detect(x, 
                                                               c("gmk","thyF","tpi","gyrB",
                                                                 "groL","hsp","dfr","16S",
                                                                 "tpi","groE","rsm", "rpoD",
                                                                 "dna"))
                                               TRUE %in% y
                                               }),]

sa_hk_genes_names <- sa_hk_genes$Gene[sa_hk_genes$No..isolates == 18]
sa_hk_genes_loci <- sa_hk_genes$GCF_000009645.1[sa_hk_genes$No..isolates == 18]

sa_hk_loci <- sa_hk_genes_loci[sa_hk_genes_loci %in% sa_counts$X]


sa_hk_counts  <- sa_counts[sa_counts$X %in% sa_hk_loci,]

hk_probs <- quantile(apply(sa_hk_counts[,-c(1,2)],2,FUN = 
                            function(x) log(max(median(x),1))), 
                    probs = c(.1,.995))

#pdf('plots/se16_comp_hk_counts.pdf', height=4, width=4)
plot(hist(apply(sa_hk_counts[,-c(1,2)],2,FUN = function(x) log(max(median(x),1))), 
          breaks = 75, plot=F),
     main = 'HK genes', xlab = 'log counts (floor 1)')
abline(v = hk_probs[1], col='red',lty=2)
abline(v = hk_probs[2], col='red',lty=2)
#dev.off()

```



## Sa zero-count genes dist 


```{r, warning=F, message=F,, fig.width=4, fig.height=4}
zc_probs <- quantile(apply(sa_counts[,-c(1,2)],2,FUN = function(x) log(max(sum(x==0),1))), 
                    probs = c(.025,.9))
#pdf('plots/se16_comp_zero_counts.pdf', height=4, width=4)
plot(hist(apply(sa_counts[,-c(1,2)],2,FUN = function(x) log(max(sum(x==0),1))),
          breaks = 50, plot=F),
     main = 'zero count genes', xlab = 'log zero counts(floor 1)')
abline(v = zc_probs[1], col='red',lty=2)
abline(v = zc_probs[2], col='red',lty=2)
#dev.off()
```


## Filer Sa data

```{r}
hk_medians <- apply(sa_hk_counts[,-c(1,2)],2,FUN = 
                            function(x) log(max(median(x),1)))
zero_counts <- apply(sa_counts[,-c(1,2)],2,FUN = 
                       function(x) log(max(sum(x==0),1)))

hk_lims <- hk_probs
zc_lims <- zc_probs
sa_counts_filtered <- sapply(c(1:length(hk_medians)), 
                             function(x){
                               if(hk_medians[x] > hk_lims[1] & 
                                  #hk_medians[x] < hk_lims[2] &
                                  #zero_counts[x] > zc_lims[1] &
                                  zero_counts[x] < zc_lims[2]){T}else{F}
                             })


```


## Heatmap of Sa showing filter


```{r, warning=F, message=F,, fig.width=12, fig.height=6}


ha <- HeatmapAnnotation(study = sa_studies, 
                        strain = sa_strains,
                        filter = sa_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = sa_ann_colors,
                                    strain = sa_st_colors,
                                    filter = c('TRUE'="white",'FALSE'="black")))

h <- Heatmap(sa_log_counts, 
             col = sa_col_fun, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
```


## Heatmap of Sa filtered


```{r, warning=F, message=F,, fig.width=12, fig.height=6}


ha <- HeatmapAnnotation(study = sa_studies[sa_counts_filtered], 
                        strain = sa_strains[sa_counts_filtered],
                        #filter = se_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = sa_ann_colors,
                                    strain = sa_st_colors#,
                                    #filter = c('TRUE'="white",'FALSE'="black")
                                    ))

h <- Heatmap(sa_log_counts[,sa_counts_filtered], 
             col = sa_col_fun, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
```


## Sa cross-mapped Se HK genes dist


```{r, warning=F, message=F,, fig.width=4, fig.height=4}


sa_cross_hk_loci <- se_hk_genes_loci[se_hk_genes_loci %in% sa_cross_counts$X]


sa_cross_hk_counts  <- sa_cross_counts[sa_cross_counts$X %in% sa_cross_hk_loci,]

hk_probs <- quantile(apply(sa_cross_hk_counts[,-c(1,2)],2,FUN = 
                            function(x) log(max(median(x),1))), 
                    probs = c(.35,.995))

#pdf('plots/se16_comp_hk_counts.pdf', height=4, width=4)
plot(hist(apply(sa_cross_hk_counts[,-c(1,2)],2,FUN = function(x) log(max(median(x),1))), 
          breaks = 75, plot=F),
     main = 'HK genes', xlab = 'log counts (floor 1)')
abline(v = hk_probs[1], col='red',lty=2)
abline(v = hk_probs[2], col='red',lty=2)
#dev.off()

```


## Sa cross-mapped zero-count dists


```{r, warning=F, message=F,, fig.width=4, fig.height=4}


zc_probs <- quantile(apply(sa_cross_counts[,-c(1,2)],2,FUN = function(x) log(max(sum(x==0),1))), 
                    probs = c(.005,.995))
#pdf('plots/se16_comp_zero_counts.pdf', height=4, width=4)
plot(hist(apply(sa_cross_counts[,-c(1,2)],2,FUN = function(x) log(max(sum(x==0),1))),
          breaks = 50, plot=F),
     main = 'zero count genes', xlab = 'log zero counts(floor 1)')
abline(v = zc_probs[1], col='red',lty=2)
abline(v = zc_probs[2], col='red',lty=2)
#dev.off()
```


## Filter Sa cross-mapped data


```{r}
hk_medians <- apply(sa_cross_hk_counts[,-c(1,2)],2,FUN = 
                            function(x) log(max(median(x),1)))
zero_counts <- apply(sa_cross_counts[,-c(1,2)],2,FUN = 
                       function(x) log(max(sum(x==0),1)))

hk_lims <- hk_probs
zc_lims <- zc_probs
sa_cross_counts_filtered <- sapply(c(1:length(hk_medians)), 
                             function(x){
                               if(hk_medians[x] > hk_lims[1] & 
                                  #hk_medians[x] < hk_lims[2] &
                                  #zero_counts[x] > zc_lims[1] &
                                  zero_counts[x] < zc_lims[2]){T}else{F}
                             })


```


## Heatmap of Sa cross-mapped data showing filter

```{r, warning=F, message=F,, fig.width=12, fig.height=6}


ha <- HeatmapAnnotation(study = sa_cross_studies, 
                        strain = sa_cross_strains,
                        filter = sa_cross_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = sax_ann_colors,
                                    strain = sax_st_colors,
                                    filter = c('TRUE'="white",'FALSE'="black")))

h <- Heatmap(sa_cross_log_counts, 
             col = sax_col_fun, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
```


## Heatmap of filtered Sa cross-mapped data 


```{r, warning=F, message=F,, fig.width=12, fig.height=6}


ha <- HeatmapAnnotation(study = sa_cross_studies[sa_cross_counts_filtered], 
                        strain = sa_cross_strains[sa_cross_counts_filtered],
                        #filter = se_counts_filtered,
                        show_legend = c(F,T), 
                        col = list( study = sax_ann_colors,
                                    strain = sax_st_colors#,
                                    #filter = c('TRUE'="white",'FALSE'="black")
                                    ))

h <- Heatmap(sa_cross_log_counts[,sa_cross_counts_filtered], 
             col = sax_col_fun, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
```


## Normalize function

```{r}
MRnorm <- function(data){
  # Description: sample-wise median of ratio normalization on columns
  # data: dataframe of gene expression counts to be normalized via DESeq2
  # returns: dataframe normalized to geometric mean of all samples
  meta <- data.frame(experiment = colnames(data))
  dds <- DESeqDataSetFromMatrix(countData =
                                       ceiling(data),
                                     colData = meta, design = ~ experiment)
  dds <- estimateSizeFactors(dds)
  counts(dds, normalized=TRUE)
}
```


## Se data normalize


```{r}
se_counts_norm  <- MRnorm(se_counts[,c(F,F,se_counts_filtered)])
rownames(se_counts_norm) <- se_counts$X

```


## Heatmap of normalized, filtered and trimmed Se data


```{r, warning=F, message=F,, fig.width=12, fig.height=6}


se_counts_norm_t <- se_counts_norm[rowSums(se_counts[,-c(1,2)]) > 50000,]
rownames(se_counts_norm_t) <- rownames(se_counts_norm[rowSums(se_counts[,-c(1,2)]) > 50000,])


se_log_counts_norm_t <- data.frame(log(se_counts_norm_t  + 1))

se_col_fun2 = colorRamp2(c(min(se_log_counts_norm_t), median(data.matrix(se_log_counts_norm_t)), max(se_log_counts_norm_t)), c("grey", "#E7D4E8", "#762A83"))




ha <- HeatmapAnnotation(study = se_studies[se_counts_filtered], 
                        strain = se_strains[se_counts_filtered],
                        #filter = se_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = se_ann_colors,
                                    strain = se_st_colors#,
                                    #filter = c('TRUE'="white",'FALSE'="black")
                                    ))


h <- Heatmap(se_log_counts_norm_t, 
             col = se_col_fun2, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             row_names_gp = gpar(fontsize=0),
             column_names_gp = gpar(fontsize=6),
             )
h
```


## Write out Se compendia


```{r}
#rownames(se_counts_norm) <- se_counts$X
se_log_counts_norm <- data.frame(log(se_counts_norm  + 1))
se_log_counts_norm_t <- data.frame(log(se_counts_norm_t  + 1))
#write.csv(se_counts_norm, 'compendia/sepi_pan_genome_clean_norm.csv', row.names = F)
#write.csv(se_log_counts_norm, 'compendia/sepi_pan_genome_clean_norm_log.csv', row.names = F)
#write.csv(se_counts_norm_t, 'compendia/sepi_pan_genome_trimmed_clean_norm.csv', row.names = F)
#write.csv(se_log_counts_norm_t, 'compendia/sepi_pan_genome_trimmed_clean_norm_log.csv', row.names = F)

```


## Log transform and write out Se data


```{r}
range <- max(se_log_counts_norm) - min(se_log_counts_norm)
range_t <- max(se_log_counts_norm_t) - min(se_log_counts_norm_t)
min(se_log_counts_norm[se_log_counts_norm > 0])

se_log_counts_norm_01 <- se_log_counts_norm / range
se_log_counts_norm_t_01 <- se_log_counts_norm_t / range_t

max(se_log_counts_norm_01)
min(se_log_counts_norm_01)

#write.csv(se_log_counts_norm_01, 'compendia/sepi_pan_genome_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(se_log_counts_norm_01, 'compendia/sepi_pan_genome_log_counts_norm_01.csv', row.names = T)

#write.csv(se_log_counts_norm_t_01, 'compendia/sepi_pan_genome_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(se_log_counts_norm_t_01, 'compendia/sepi_pan_genome_trimmed_log_counts_norm_01.csv', row.names = T)
```


## Sa data normalize


```{r}
sa_counts_norm  <- MRnorm(sa_counts[,c(F,F,sa_counts_filtered)])
rownames(sa_counts_norm) <- sa_counts$X
```


## Heatmap of Sa data filtered, normalized, trimmed


```{r, warning=F, message=F,, fig.width=12, fig.height=6}

sa_counts_norm_t <- sa_counts_norm[rowSums(sa_counts[,-c(1,2)]) > 500000,]
dim(sa_counts_norm_t)

sa_log_counts_norm_t <- data.frame(log(sa_counts_norm_t  + 1))

sa_col_fun2 = colorRamp2(c(min(sa_log_counts_norm_t), median(data.matrix(sa_log_counts_norm_t)), max(sa_log_counts_norm_t)), c("grey", "#E7D4E8", "#762A83"))


ha <- HeatmapAnnotation(study = sa_studies[sa_counts_filtered], 
                        strain = sa_strains[sa_counts_filtered], 
                        show_legend = F, col = list( study = sa_ann_colors,
                                                     strain = sa_st_colors))

h <- Heatmap(sa_log_counts_norm_t, 
             col = sa_col_fun2, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(6,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             row_names_gp = gpar(fontsize=0)
             )
h
```


## Write out Sa compendia


```{r}
sa_log_counts_norm <- data.frame(log(sa_counts_norm  + 1))
sa_log_counts_norm_t <- data.frame(log(sa_counts_norm_t  + 1))
#write.csv(sa_counts_norm, 'compendia/saur_pan_genome_clean_norm.csv', row.names = F)
#write.csv(sa_log_counts_norm, 'compendia/saur_pan_genome_clean_norm_log.csv', row.names = F)
#write.csv(sa_counts_norm_t, 'compendia/saur_pan_genome_trimmed_clean_norm.csv', row.names = F)
#write.csv(sa_log_counts_norm_t, 'compendia/saur_pan_genome_trimmed_clean_norm_log.csv', row.names = F)

```


## Log transform and write out Sa comps


```{r}
range <- max(sa_log_counts_norm) - min(sa_log_counts_norm)
range_t <- max(sa_log_counts_norm_t) - min(sa_log_counts_norm_t)
min(sa_log_counts_norm[sa_log_counts_norm > 0])

sa_log_counts_norm_01 <- sa_log_counts_norm / range
sa_log_counts_norm_t_01 <- sa_log_counts_norm_t / range_t

max(sa_log_counts_norm_01)
min(sa_log_counts_norm_01)

#write.csv(sa_log_counts_norm_01, 'compendia/saur_pan_genome_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_log_counts_norm_01, 'compendia/saur_pan_genome_log_counts_norm_01.csv', row.names = T)

#write.csv(sa_log_counts_norm_t_01, 'compendia/saur_pan_genome_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_log_counts_norm_t_01, 'compendia/saur_pan_genome_trimmed_log_counts_norm_01.csv', row.names = T)
```



## Normalize cross-mapped data


```{r}
sa_cross_counts_norm  <- MRnorm(sa_cross_counts[,c(F,F,sa_cross_counts_filtered)])
rownames(sa_cross_counts_norm) <- sa_cross_counts$X
```


## Heatmap of Sa cross-mapped data filterend, norm, trimmed


```{r, warning=F, message=F,, fig.width=12, fig.height=6}

sa_cross_counts_norm_t <- sa_cross_counts_norm[rowSums(sa_cross_counts[,-c(1,2)]) > 50000,]
rownames(sa_cross_counts_norm_t) <- rownames(sa_cross_counts_norm[rowSums(sa_cross_counts[,-c(1,2)]) > 50000,])
sa_cross_log_counts_norm_t <- data.frame(log(sa_cross_counts_norm_t  + 1))

sax_col_fun2 = colorRamp2(c(min(sa_cross_log_counts_norm_t), median(data.matrix(sa_cross_log_counts_norm_t)), max(sa_cross_log_counts_norm_t)), c("grey", "#E7D4E8", "#762A83"))



ha <- HeatmapAnnotation(study = sa_cross_studies[sa_cross_counts_filtered], 
                        strain = sa_cross_strains[sa_cross_counts_filtered], 
                        show_legend = F, col = list( study = sax_ann_colors,
                                                     strain = sax_st_colors))

h <- Heatmap(sa_cross_log_counts_norm_t, 
             col = sax_col_fun2, 
             top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             row_names_gp = gpar(fontsize=0)
             )
h
```


## Write out cross mapped data


```{r}
sa_cross_log_counts_norm <- data.frame(log(sa_cross_counts_norm  + 1))
sa_cross_log_counts_norm_t <- data.frame(log(sa_cross_counts_norm_t  + 1))
#write.csv(sa_cross_counts_norm, 'compendia/saur_pan_genome_CROSS_clean_norm.csv', row.names = F)
#write.csv(sa_cross_log_counts_norm, 'compendia/saur_pan_genome_CROSS_clean_norm_log.csv', row.names = F)
#write.csv(sa_cross_counts_norm_t, 'compendia/saur_pan_genome_CROSS_trimmed_clean_norm.csv', row.names = F)
#write.csv(sa_cross_log_counts_norm_t, 'compendia/saur_pan_genome_CROSS_trimmed_clean_norm_log.csv', row.names = F)

```


## Log transform and write out cross mapped data


```{r}
range <- max(sa_cross_log_counts_norm) - min(sa_cross_log_counts_norm)
range_t <- max(sa_cross_log_counts_norm_t) - min(sa_cross_log_counts_norm_t)
min(sa_cross_log_counts_norm[sa_cross_log_counts_norm > 0])

sa_cross_log_counts_norm_01 <- sa_cross_log_counts_norm / range
sa_cross_log_counts_norm_t_01 <- sa_cross_log_counts_norm_t / range_t

max(sa_cross_log_counts_norm_01)
min(sa_cross_log_counts_norm_01)

#write.csv(sa_cross_log_counts_norm_01, 'compendia/saur_pan_genome_CROSS_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_cross_log_counts_norm_01, 'compendia/saur_pan_genome_CROSS_log_counts_norm_01.csv', row.names = T)

#write.csv(sa_cross_log_counts_norm_t_01, 'compendia/saur_pan_genome_CROSS_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_cross_log_counts_norm_t_01, 'compendia/saur_pan_genome_CROSS_trimmed_log_counts_norm_01.csv', row.names = T)
```



## Re-order Se by RAND to match Se


```{r}
set.seed(22)
rand_order <-  sample(rownames(se_log_counts_norm_01), nrow(se_log_counts_norm_01), replace = F)

se_log_counts_norm_01_rand <- se_log_counts_norm_01[rand_order ,]
rownames(se_log_counts_norm_01_rand) <- rownames(se_log_counts_norm_01)
#write.csv(se_log_counts_norm_01_rand, 'compendia/sepi_pan_genome_RAND_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(se_log_counts_norm_01_rand, 'compendia/sepi_pan_genome_RAND_log_counts_norm_01.csv', row.names = T)



rand_order_t <- sample(rownames(se_log_counts_norm_t_01), nrow(se_log_counts_norm_t_01), replace = F)
se_log_counts_norm_t_01_rand <- se_log_counts_norm_t_01[rand_order_t,]
se_log_counts_norm_t_01_rand[is.na(se_log_counts_norm_t_01_rand)] <- 0
rownames(se_log_counts_norm_t_01_rand) <- rownames(se_log_counts_norm_t_01)
#write.csv(se_log_counts_norm_t_01_rand, 'compendia/sepi_pan_genome_RAND_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(se_log_counts_norm_t_01_rand, 'compendia/sepi_pan_genome_RAND_trimmed_log_counts_norm_01.csv', row.names = T)

```



## Re-order Sa cross-mapped by RAND to match Se


```{r}

sa_cross_log_counts_norm_01_rand <- sa_cross_log_counts_norm_01[rand_order ,]
rownames(sa_cross_log_counts_norm_01_rand) <- rownames(se_log_counts_norm_01)

#write.csv(sa_cross_log_counts_norm_01_rand, 'compendia/saur_pan_genome_CROSS_RAND_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_cross_log_counts_norm_01_rand, 'compendia/saur_pan_genome_CROSS_RAND_log_counts_norm_01.csv', row.names = T)



sa_cross_log_counts_norm_t_01_rand <- sa_cross_log_counts_norm_t_01[rand_order_t ,]
rownames(sa_cross_log_counts_norm_t_01_rand) <- rownames(se_log_counts_norm_t_01)
#write.csv(sa_cross_log_counts_norm_t_01_rand, 'compendia/saur_pan_genome_CROSS_RAND_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_cross_log_counts_norm_t_01_rand, 'compendia/saur_pan_genome_CROSS_RAND_trimmed_log_counts_norm_01.csv', row.names = T)

```





## Re-order Sa by BLAST to match Se


```{r}
pan_map <- read.csv('se_sa_pan_map.csv', stringsAsFactors = F)
pan_map <- pan_map[order(pan_map$hit_evalue),]
rownames(pan_map) <- make.unique(pan_map$se_ID)

se_gene_order <- rownames(se_log_counts_norm_01)
se_gene_order_sa <- pan_map[se_gene_order, "sa_ID"]

sa_blast_map <- sa_log_counts_norm_01[se_gene_order_sa,]
rownames(sa_blast_map) <- se_gene_order
sa_blast_map[is.na(sa_blast_map)] <- 0

#write.csv(sa_blast_map, 'compendia/saur_pan_genome_BLAST_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_blast_map, 'compendia/saur_pan_genome_BLAST_log_counts_norm_01.csv', row.names = T)
```


```{r}
m <- melt(sa_blast_map)
g <- ggplot(m, aes(x = value + (min(m$value[m$value > 0]) / 2))) +
  geom_histogram(aes(y=after_stat(count))) +
  #geom_density() +
  scale_x_log10() +
  scale_y_log10() +
  theme_prism()
g
```


```{r}
se_gene_order_t <- rownames(se_log_counts_norm_t_01)
se_gene_order_sa_t <- pan_map[se_gene_order_t, "sa_ID"]

sa_blast_map_t <- sa_log_counts_norm_t_01[se_gene_order_sa_t,]
rownames(sa_blast_map_t) <- se_gene_order_t
sa_blast_map_t[is.na(sa_blast_map_t)] <- 0

#write.csv(sa_blast_map_t, 'compendia/saur_pan_genome_BLAST_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_blast_map_t, 'compendia/saur_pan_genome_BLAST_trimmed_log_counts_norm_01.csv', row.names = T)
```


```{r}
m <- melt(sa_blast_map_t)
g <- ggplot(m, aes(x = value + (min(m$value[m$value > 0]) / 2))) +
  geom_histogram(aes(y=after_stat(count))) +
  #geom_density() +
  scale_x_log10() +
  scale_y_log10() +
  theme_prism()
g
```



```{r, warning=F, message=F,, fig.width=12, fig.height=6}

sa_col_fun3 = colorRamp2(c(min(sa_blast_map_t), median(data.matrix(sa_blast_map_t)), max(sa_blast_map_t)), c("grey", "#E7D4E8", "#762A83"))

ha <- HeatmapAnnotation(study = sa_studies[sa_counts_filtered], 
                        strain = sa_strains[sa_counts_filtered],
                        #filter = se_counts_filtered,
                        show_legend = c(F,T,T), 
                        col = list( study = sa_ann_colors,
                                    strain = sa_st_colors#,
                                    #filter = c('TRUE'="white",'FALSE'="black")
                                    ))

h <- Heatmap(sa_blast_map_t, 
             col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
h
#png('plots/Sa_BLAST_noclust.#png', height=6, width=12, unit = "in", res = 150)
#h
#dev.off()
```




## Re-order Sa  by RAND to match Se

```{r}


rand_order_sa <-  sample(rownames(sa_log_counts_norm_01), length(se_gene_order_sa), replace = T)

sa_log_counts_norm_01_rand <- sa_log_counts_norm_01[rand_order_sa ,]
rownames(sa_log_counts_norm_01_rand) <- rownames(se_log_counts_norm_01)

sum(!is.na(rowSums(sa_log_counts_norm_01_rand)))

sa_log_counts_norm_01_rand[is.na(sa_log_counts_norm_01_rand)] <- 0

#write.csv(sa_log_counts_norm_01_rand, 'compendia/saur_pan_genome_RAND_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_log_counts_norm_01_rand, 'compendia/saur_pan_genome_RAND_log_counts_norm_01.csv', row.names = T)
```



```{r}

rand_order_sa_t <- rand_order_sa[rowSums(se_counts_norm) > 500]
sa_log_counts_norm_t_01_rand <- sa_log_counts_norm_t_01[rand_order_sa_t,]
sa_log_counts_norm_t_01_rand[is.na(sa_log_counts_norm_t_01_rand)] <- 0
#rownames(sa_log_counts_norm_t_01_rand) <- se_gene_order_t

#write.csv(sa_log_counts_norm_t_01_rand, 'compendia/saur_pan_genome_RAND_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_log_counts_norm_t_01_rand, 'compendia/saur_pan_genome_RAND_trimmed_log_counts_norm_01.csv', row.names = T)
```



```{r, warning=F, message=F,, fig.width=12, fig.height=6}





h <- Heatmap(sa_log_counts_norm_t_01_rand, 
             col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
            cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
#png('plots/Sa_RAND_noclust.#png', height=6, width=12, unit = "in", res = 150)
h
#dev.off()
```



## Re-order Sa by BLAST with RAND to match Se


```{r}

se_gene_order_sa_bg_rand <- se_gene_order_sa
se_gene_order_sa_bg_rand[(se_gene_order_sa == "") | is.na(se_gene_order_sa)] <- rand_order_sa[(se_gene_order_sa == "") | is.na(se_gene_order_sa)]

sa_blast_map_randfill <- sa_log_counts_norm_01[se_gene_order_sa_bg_rand,]
rownames(sa_blast_map_randfill) <- se_gene_order
sum(!is.na(rowSums(sa_blast_map_randfill, na.rm=T)))

sa_blast_map_randfill[is.na(sa_blast_map_randfill)] <- 0

#write.csv(sa_blast_map_randfill, 'compendia/saur_pan_genome_BLAST_with_rand_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_blast_map_randfill, 'compendia/saur_pan_genome_BLAST_with_rand_log_counts_norm_01.csv', row.names = T)
```



```{r}
m <- melt(sa_blast_map_randfill)
g <- ggplot(m, aes(x = value + (min(m$value[m$value > 0]) / 2))) +
  geom_histogram(aes(y=after_stat(count))) +
  #geom_density() +
  scale_x_log10() +
  scale_y_log10() +
  theme_prism()
g
```




```{r}
se_gene_order_sa_bg_rand_t <- se_gene_order_sa_t
se_gene_order_sa_bg_rand_t[(se_gene_order_sa_t == "") | is.na(se_gene_order_sa_t)] <- rand_order_sa_t[(se_gene_order_sa_t == "") | is.na(se_gene_order_sa_t)]


sa_blast_map_randfill_t <- sa_log_counts_norm_t_01[se_gene_order_sa_bg_rand_t,]
#rownames(sa_blast_map_randfill_t) <- se_gene_order_t
sum(!is.na(rowSums(sa_blast_map_randfill_t, na.rm=T)))

sa_blast_map_randfill_t[is.na(sa_blast_map_randfill_t)] <- 0

#write.csv(sa_blast_map_randfill_t, 'compendia/saur_pan_genome_BLAST_with_rand_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_blast_map_randfill_t, 'compendia/saur_pan_genome_BLAST_with_rand_trimmed_log_counts_norm_01.csv', row.names = T)
```


```{r}
m <- melt(sa_blast_map_randfill_t)
g <- ggplot(m, aes(x = value + (min(m$value[m$value > 0]) / 2))) +
  geom_histogram(aes(y=after_stat(count))) +
  #geom_density() +
  scale_x_log10() +
  scale_y_log10() +
  theme_prism()
g
```



```{r, warning=F, message=F,, fig.width=12, fig.height=6}


h <- Heatmap(sa_blast_map_t, 
             col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
#png('plots/Sa_BLASTwRand_noclust.#png', height=6, width=12, unit = "in", res = 150)
h
#dev.off()
```






## Re-order Sa by FOLDSEEK to match Se


```{r, include=F}
fs_map <- read.csv('saur_sepi_aln_g.m8', stringsAsFactors = F, sep = '\t',
                   header = F)

fs_map_best_hits <- fs_map[order(fs_map$V11),]
rownames(fs_map_best_hits) <- make.unique(fs_map_best_hits$V2)
fs_map_best_hits <- fs_map_best_hits[rownames(fs_map_best_hits) %in% fs_map_best_hits$V2,]

rownames(fs_map_best_hits) <- make.unique(fs_map_best_hits$V1)

firsts <- fs_map_best_hits[rownames(fs_map_best_hits) %in% fs_map_best_hits$V1,]
n_firsts <- fs_map_best_hits[!rownames(fs_map_best_hits) %in% fs_map_best_hits$V1,]


fs_map_best_hits_rev <- fs_map[order(fs_map$V11),]
rownames(fs_map_best_hits_rev) <- make.unique(fs_map_best_hits_rev$V1)
fs_map_best_hits_rev <- fs_map_best_hits_rev[rownames(fs_map_best_hits_rev) %in% fs_map_best_hits_rev$V1,]

length(unique(fs_map_best_hits_rev$V1))
length(unique(fs_map_best_hits_rev$V2))

secs <- fs_map_best_hits_rev[fs_map_best_hits_rev$V2 %in% n_firsts$V2,]
rownames(secs) <- make.unique(secs$V2)

sum(rownames(secs) %in% secs$V2)

fs_map_out <- rbind(firsts, secs, n_firsts[!n_firsts$V2 %in% secs$V2,])


sa_gene_loci$name_id[is.na(sa_gene_loci$name_id)] <- ""
rownames(sa_gene_loci) <- make.unique(sa_gene_loci$name_id)

se_gene_loci$name_id[is.na(se_gene_loci$name_id)] <- ""
rownames(se_gene_loci) <- make.unique(se_gene_loci$name_id)


fs_map_out$sa_ID <- sa_gene_loci[fs_map_out$V1, "locus_id"]
fs_map_out$se_ID <- se_gene_loci[fs_map_out$V2, "locus_id"]

fs_map_out$sa_ID[is.na(fs_map_out$sa_ID)] <- ""
fs_map_out$se_ID[is.na(fs_map_out$se_ID)] <- ""

rownames(fs_map_out) <- make.unique(fs_map_out$se_ID)

sum(se_gene_order %in% rownames(fs_map_out))

se_gene_order_sa_fs <- fs_map_out[se_gene_order, "sa_ID"]

sa_fs_map <- sa_log_counts_norm_01[se_gene_order_sa_fs,]
rownames(sa_fs_map) <- se_gene_order
sum(!is.na(rowSums(sa_fs_map)))

sa_fs_map[is.na(sa_fs_map)] <- 0

#write.csv(sa_fs_map, 'compendia/saur_pan_genome_FLDSK_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map, 'compendia/saur_pan_genome_FLDSK_log_counts_norm_01.csv', row.names = T)
```

```{r}
#write.csv(fs_map_out, "foldseek_mapping_out.csv", row.names=T)
```


```{r}
fs_map_out <- read.csv("foldseek_mapping_out.csv", stringsAsFactors = F, row.names = 1)



#se_gene_order <- rownames(se_log_counts_norm_01)
se_gene_order_sa <- fs_map_out[se_gene_order, "sa_ID"]
se_gene_order_sa_fs <- fs_map_out[se_gene_order, "sa_ID"]

sa_fs_map <- sa_log_counts_norm_01[se_gene_order_sa,]
rownames(sa_fs_map) <- se_gene_order
sa_fs_map[is.na(sa_fs_map)] <- 0

#write.csv(sa_fs_map, 'compendia/saur_pan_genome_FLDSK_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map, 'compendia/saur_pan_genome_FLDSK_log_counts_norm_01.csv', row.names = T)
```


```{r}

se_gene_order_sa_t_fs <- fs_map_out[se_gene_order_t, "sa_ID"]

sa_fs_map_t <- sa_log_counts_norm_01[se_gene_order_sa_t_fs,]
rownames(sa_fs_map_t) <- se_gene_order_t
sa_fs_map_t[is.na(sa_fs_map_t)] <- 0

#write.csv(sa_fs_map_t, 'compendia/saur_pan_genome_FLDSK_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map_t, 'compendia/saur_pan_genome_FLDSK_trimmed_log_counts_norm_01.csv', row.names = T)
```


```{r, warning=F, message=F,, fig.width=12, fig.height=6}



h <- Heatmap(sa_blast_map_t, 
             col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
#png('plots/Sa_FLDSK_noclust.#png', height=6, width=12, unit = "in", res = 150)
h
#dev.off()
```

## Re-order Sa by FOLDSEEK with RAND to match Se


```{r}


se_gene_order_sa_fs[(se_gene_order_sa_fs == "") | is.na(se_gene_order_sa_fs)] <- rand_order_sa[(se_gene_order_sa_fs == "") | is.na(se_gene_order_sa_fs)]

sa_fs_map_randfill <- sa_log_counts_norm_01[se_gene_order_sa_fs,]
rownames(sa_fs_map_randfill) <- se_gene_order
sum(!is.na(rowSums(sa_fs_map_randfill)))

sa_fs_map_randfill[is.na(sa_fs_map_randfill)] <- 0

#write.csv(sa_fs_map_randfill, 'compendia/saur_pan_genome_FLDSK_with_rand_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map_randfill, 'compendia/saur_pan_genome_FLDSK_with_rand_log_counts_norm_01.csv', row.names = T)
```



```{r}


se_gene_order_sa_t_fs[(se_gene_order_sa_t_fs == "") | is.na(se_gene_order_sa_t_fs)] <- rand_order_sa_t[(se_gene_order_sa_t_fs == "") | is.na(se_gene_order_sa_t_fs)]

sa_fs_map_t_randfill <- sa_log_counts_norm_01[se_gene_order_sa_t_fs,]
rownames(sa_fs_map_t_randfill) <- se_gene_order_t
sum(!is.na(rowSums(sa_fs_map_t_randfill)))

sa_fs_map_t_randfill[is.na(sa_fs_map_t_randfill)] <- 0

#write.csv(sa_fs_map_t_randfill, 'compendia/saur_pan_genome_FLDSK_with_rand_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map_t_randfill, 'compendia/saur_pan_genome_FLDSK_with_rand_trimmed_log_counts_norm_01.csv', row.names = T)
```


```{r, warning=F, message=F,, fig.width=12, fig.height=6}




h <- Heatmap(sa_fs_map_t_randfill, col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
#png('plots/Sa_FLDSKwRand_noclust.#png', height=6, width=12, unit = "in", res = 150)
h
#dev.off()
```

## Re-order Sa by FOLDSEEK-EXHUASTIVE to match Se


```{r, include=F}

fs_map <- read.csv('saur_sepi_aln_g_exh.m8', stringsAsFactors = F, sep = '\t',
                   header = F)

fs_map_best_hits <- fs_map[order(fs_map$V11),]
rownames(fs_map_best_hits) <- make.unique(fs_map_best_hits$V2)
fs_map_best_hits <- fs_map_best_hits[rownames(fs_map_best_hits) %in% fs_map_best_hits$V2,]

rownames(fs_map_best_hits) <- make.unique(fs_map_best_hits$V1)

firsts <- fs_map_best_hits[rownames(fs_map_best_hits) %in% fs_map_best_hits$V1,]
n_firsts <- fs_map_best_hits[!rownames(fs_map_best_hits) %in% fs_map_best_hits$V1,]

length(unique(fs_map_best_hits$V1))
length(unique(fs_map_best_hits$V2))

fs_map_best_hits_rev <- fs_map[order(fs_map$V11),]
rownames(fs_map_best_hits_rev) <- make.unique(fs_map_best_hits_rev$V1)
fs_map_best_hits_rev <- fs_map_best_hits_rev[rownames(fs_map_best_hits_rev) %in% fs_map_best_hits_rev$V1,]

length(unique(fs_map_best_hits_rev$V1))
length(unique(fs_map_best_hits_rev$V2))

secs <- fs_map_best_hits_rev[fs_map_best_hits_rev$V2 %in% n_firsts$V2,]
rownames(secs) <- make.unique(secs$V2)

sum(rownames(secs) %in% secs$V2)

fs_map_out <- rbind(firsts, secs, n_firsts[!n_firsts$V2 %in% secs$V2,])


length(unique(fs_map_out$V1))
length(unique(fs_map_out$V2))

sa_gene_loci$name_id[is.na(sa_gene_loci$name_id)] <- ""
rownames(sa_gene_loci) <- make.unique(sa_gene_loci$name_id)

se_gene_loci$name_id[is.na(se_gene_loci$name_id)] <- ""
rownames(se_gene_loci) <- make.unique(se_gene_loci$name_id)


fs_map_out$sa_ID <- sa_gene_loci[fs_map_out$V1, "locus_id"]
fs_map_out$se_ID <- se_gene_loci[fs_map_out$V2, "locus_id"]

fs_map_out$sa_ID[is.na(fs_map_out$sa_ID)] <- ""
fs_map_out$se_ID[is.na(fs_map_out$se_ID)] <- ""

rownames(fs_map_out) <- make.unique(fs_map_out$se_ID)

sum(se_gene_order %in% rownames(fs_map_out))

se_gene_order_sa_fsexh <- fs_map_out[se_gene_order, "sa_ID"]

sa_fsexh_map <- sa_log_counts_norm_01[se_gene_order_sa_fsexh,]
rownames(sa_fsexh_map) <- se_gene_order
sum(!is.na(rowSums(sa_fsexh_map)))

sa_fsexh_map[is.na(sa_fsexh_map)] <- 0

#write.csv(sa_fsexh_map, 'compendia/saur_pan_genome_FLDSKEXH_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fsexh_map, 'compendia/saur_pan_genome_FLDSKEXH_log_counts_norm_01.csv', row.names = T)

```

```{r}
#write.csv(fs_map_out, "foldseek_exhaustive_mapping_out.csv", row.names=T)
```


```{r}
fs_map_out <- read.csv("foldseek_exhaustive_mapping_out.csv", stringsAsFactors = F, row.names = 1)



#se_gene_order <- rownames(se_log_counts_norm_01)
se_gene_order_sa <- fs_map_out[se_gene_order, "sa_ID"]
se_gene_order_sa_fs <- fs_map_out[se_gene_order, "sa_ID"]

sa_fs_map <- sa_log_counts_norm_01[se_gene_order_sa,]
rownames(sa_fs_map) <- se_gene_order
sa_fs_map[is.na(sa_fs_map)] <- 0

#write.csv(sa_fs_map, 'compendia/saur_pan_genome_FLDSKEXH_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map, 'compendia/saur_pan_genome_FLDSKEXH_log_counts_norm_01.csv', row.names = T)
```


```{r}

se_gene_order_sa_t_fsexh <- fs_map_out[se_gene_order_t, "sa_ID"]

sa_fsexh_map_t <- sa_log_counts_norm_01[se_gene_order_sa_t_fsexh,]
rownames(sa_fsexh_map_t) <- se_gene_order_t
sa_fsexh_map_t[is.na(sa_fsexh_map_t)] <- 0

#write.csv(sa_fsexh_map_t, 'compendia/saur_pan_genome_FLDSKEXH_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fsexh_map_t, 'compendia/saur_pan_genome_FLDSKEXH_trimmed_log_counts_norm_01.csv', row.names = T)
```


```{r, warning=F, message=F,, fig.width=12, fig.height=6}



h <- Heatmap(sa_fsexh_map_t, 
             col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
#png('plots/Sa_FLDSKEXH_noclust.#png', height=6, width=12, unit = "in", res = 150)
h
#dev.off()
```

## Re-order Sa by FOLDSEEK-EXHUASTIVE with RAND to match Se


```{r}



se_gene_order_sa_fs[(se_gene_order_sa_fs == "") | is.na(se_gene_order_sa_fs)] <- rand_order_sa[(se_gene_order_sa_fs == "") | is.na(se_gene_order_sa_fs)]

sa_fs_map_randfill <- sa_log_counts_norm_01[se_gene_order_sa_fs,]
rownames(sa_fs_map_randfill) <- se_gene_order
sum(!is.na(rowSums(sa_fs_map_randfill)))

sa_fs_map_randfill[is.na(sa_fs_map_randfill)] <- 0

#write.csv(sa_fs_map_randfill, 'compendia/saur_pan_genome_FLDSKEXH_with_rand_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fs_map_randfill, 'compendia/saur_pan_genome_FLDSKEXH_with_rand_log_counts_norm_01.csv', row.names = T)
```




```{r}
se_gene_order_sa_t_fs[(se_gene_order_sa_t_fs == "") | is.na(se_gene_order_sa_t_fs)] <- rand_order_sa_t[(se_gene_order_sa_t_fs == "") | is.na(se_gene_order_sa_t_fs)]

sa_fs_map_t_randfill <- sa_log_counts_norm_01[se_gene_order_sa_t_fs,]
rownames(sa_fs_map_t_randfill) <- se_gene_order_t
sum(!is.na(rowSums(sa_fs_map_t_randfill)))

sa_fs_map_t_randfill[is.na(sa_fs_map_t_randfill)] <- 0

#write.csv(sa_fsexh_map_randfill_t, 'compendia/saur_pan_genome_FLDSKEXH_with_rand_trimmed_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_fsexh_map_randfill_t, 'compendia/saur_pan_genome_FLDSKEXH_with_rand_trimmed_log_counts_norm_01.csv', row.names = T)
```


```{r, warning=F, message=F,, fig.width=12, fig.height=6}



h <- Heatmap(sa_fs_map_t_randfill, col = sa_col_fun3, 
            top_annotation = ha,
             height = unit(8,"cm"),
             width = unit(24,"cm"),
             cluster_columns = F,
             cluster_rows = F,
             heatmap_legend_param = list(
               title = 'log counts',
             legend_height = unit(8,"cm"),
             title_position = "leftcenter-rot"
             ),
             column_names_gp = gpar(fontsize=6),
             show_row_names = F,
             show_column_names = F
             )
#png('plots/Sa_FLDSKEXHwRand_noclust.#png', height=6, width=12, unit = "in", res = 150)
h
#dev.off()
```



## Re-order Sa by RAND to match Sa


```{r}
rand_order <-  sample(rownames(sa_log_counts_norm_01), nrow(sa_log_counts_norm_01), replace = F)

sa_log_counts_norm_01_rand <- sa_log_counts_norm_01[rand_order ,]
rownames(sa_log_counts_norm_01_rand) <- rownames(sa_log_counts_norm_01)
#write.csv(sa_log_counts_norm_01_rand, 'compendia/saur_pan_genome_RANDshuff_log_counts_norm_01_nonames.csv', row.names = F)
#write.csv(sa_log_counts_norm_01_rand, 'compendia/saur_pan_genome_RANDshuff_log_counts_norm_01.csv', row.names = T)

```


